A hierarchical approach for modelling X-ray beamlines. Application to a coherent beamline.

### 2019-04-03 paper-hierarchical.tex
version sent to coauthors

\begin{abstract}
We consider different approaches to simulate a modern X-ray beamline. Several methodologies with increasing complexity are applied to discuss the relevant parameters that quantify the beamline performances. Parameters such as flux, beam size at the image and coherence properties are obtained from simple analytical calculations to sophisticated computer simulations using ray tracing and waveoptics techniques. A last generation x-ray nanofocusing beamline for coherent  applications (ID16A at the ESRF) has been chosen  to study in detail the issues 
related to highly demagnifying sychrotron sources and exploiting the beam coherence. The performance of the beamline is studied for two storages rings: the old ESRF (emittance 4000 pm) and the new ESRF-EBS (emittance 150 pm). In addition to traditional results in term of flux, intensity distribution and focal size, an innovative study on the partial coherence properties based on propagation of coherent modes is presented. The different algorithms and methodologies are implemented in the software suite OASYS \cite{codeOASYS}. They are discussed with emphasis in benefits and limitations of each one. 
\end{abstract}

@@

Many storage ring based x-ray synchrotron facilities are building or planning upgrades to increase brilliance and coherent flux by one to three orders of magnitude.  The first upgrade of a large facility will be the EBS (Extremely Brilliant Source) \cite{orangebook} at the European Synchrotron Radiation Facility (ESRF), aiming to build a storage ring of 150 pm emittance to significantly boost the associated x-ray coherence. Applications exploiting beam coherence such as such as x-ray photon correlation spectroscopy, coherent diffraction imaging, propagation-based phase contrast imaging and ptychography will strongly benefit from this update.

Accurate calculation and quantitative evaluation of the parameters related to x-ray optics is of paramount importance for designing, building and exploiting the new beamlines. Every modern beamline follows a procedure of conceptual design and then technical design where detailed simulations are essential. The calculations allow the designer to test the design parameters in a virtual computer environment where pros and cons can be studied quantitatively. The selection to the best performant optics configuration requires the detailed simulation of the optics imperfections, that in many cases are the limiting factor of the beamline optical configuration. Deformation in the optical elements due to heat load need to be controlled, therefore optics simulations have to include results from engineering modeling of the thermal deformation associated to mechanical issues usually dome using finite eelement analysis. Traditionally the optics calculations for synchrotron beamlines are done using ray tracing techniques. They model the x-ray beam as a collection of different rays that are incoherent one to another. This technique is still essential for obtaining information on aberrations, flux propagation, monochromator bandwidth, etc. However, the high coherence of the new sources make necessary to complement the ray tracing methods with models including effects of diffraction and scattering from a coherent beam. A fully coherent beam is still a dream for the new generation storage rings. In spite of a high increate of the coherent fraction of the new beams, the coherent fraction at tipically hard x-ray photon energies is still at few percent for the new sources. This makes still important a ray tracing analysis that does not become obsolete, but has to be complemented with an analysis based on wave optics. Because of the non perfect coherence the wave optics analysis becomes non trivial and can be done with different levels of approximation, as discussed in this work. In addition, wave optics simulations are usually expensive from the computer point of view.

For illustrating the applicability of a battery of methods and solutions for beamline optics simulations we have chosen the ESRF beamline ID16A. This beamline provides a high-brilliance beam focused to a few nanometer. It shows characteristics increasingly searched in modern beamlines: extremely high demagnification (sub-micrometer beam) and high coherence. It combines coherent imaging techniques and X-ray fluorescence microscopy to perform quantitative 3D characterization of the morphology and the elemental composition of specimens in their native state \todo{complete ID16A description}.

X-ray beamlines are very particular optical systems. Some implications of exploiting low wavelengths such as the necessity of using grazing incidence reflectors and the use of perfect crystals as typical hard x-ray monochromators make useless, or at least incomplete, the application of common commercial software packages for the study of synchrotron beamlines. For this reason many optics tools have been developed by the synchrotron community. Some of them have a long history like SHADOW (more than 40 years old) or SRW (almost 30), for mentioning only codes used in this work. These historical tools together with new tools designed ad-hoc packages have been agglutinated in a user friendly software OASYS. Many of the calculations presented here have been performed directly in OASYS or with script created with the help of OASYS.

The paper is structures as follows: we first introduce in Section~\ref{Beamline description} the main parameters of the beamline and make the selection of the particular choice of the parameters used in the calculations. Section~\label{Simulations} is the central part of the paper, and describe the different methods used to analyse the beamline with increasing complexity, setting a hierarchy of methodologies that can be exported to the simulations of any beamline. In Section \ref{level0} we calculate with simple methods approximated values of focal size, flux and coherent fraction. Then accuracy is improved using geometrical methods in Section \ref{level1} that give reasonable results as far as some corrections are considered because of the  mentioned characteristics of high demagnification and high coherence of the studied beamline. The wave optics section~]ref{level2} also include different models whose complexity increases when the results expected are more and mode demanding. The simulations methods are presented in a hierarchical approach:  they provide more and better results but the price to pay is to deal with more complex and computationally expensive tools.

 %-------------------------------------------------------------------------
 % Manuel & Rafael
 %-------------------------------------------------------------------------

@@

\resizebox{\textwidth}{!}{%
\begin{tabular}{lcccccc}
name  & $\lambda_u$ [mm] & $L$ [m] & N & $K$ (11200 eV) & $K$ (17225 eV) & $K$ (33600 eV) \\ 
\input{tableUndulator.txt}
%      & IVU26 & 26.00           & 2.50    & 1      & 0.616 (n=1)    & 1.625 (n=3)    & 0.616 (n=3)    \\
% ESRF & U18.3 & 18.30           & 1.40    & 2      & 1.715 (n=1)    & 0.445 (n=1)    & 1.715 (n=3)    \\
%      & U22.4 & 22.40           & 1.40    & 1      & 0.873 (n=1)    & 1.840 (n=1)    & 0.873 (n=1)    \\ \hline
%      & IVU26 & 26.00           & 2.50    & 1      & 0.590 (n=1)    & 1.606 (n=1)    & 0.590 (n=3)    \\
% EBS  & U18.3 & 18.30           & 1.40    & 2      & 1.156 (n=1)    & 0.411 (n=1)    & 1.156 (n=1)    \\
%      & U22.4 & 22.40           & 1.40    & 1      & 0.852 (n=1)    & 1.821 (n=3)    & 0.852 (n=1)    \\ \hline
\end{tabular}%
}
\end{table}

@@

\begin{table}\label{tab:Positions}
    \caption{Position of the main ID16A optical elements with respect to the source (undulator centre). It also shows the values for horizontal and vertical demagnification $M^{-1}=p/q$, where $p$ and $q$ are the the distances object-optics and optics-image, respectively.}
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{rcccccc}
    -                      & Source & Multilayer & Slit  & KB (v) & KB (h) & Focal plane \\ 
    Position {[}m{]}:      & 0.00   & 28.30      & 40.00 & 184.90 & 184.95 & 185.00      \\
    Horizontal $M^{-1}$:   & -      & 2.42       & -     & -      & 2899   & -           \\
    Vertical $M^{-1}$:     & -      & -          & -     & \inblue{1849}   & -      & -           \\ 
    \end{tabular}%
    }%
\end{table}

Simulations were done using the two different storage ring lattices mentioned before: the (present) ESRF DBA where the undulators are in a high-$\beta$ straight section (larger horizontal emittance) and the new EBS (lower horizontal emittance) straight sections. The parameters for the current and the new lattice are shown in Tables \ref{tab:TableSources} and \ref{tab:eBeam}.

@@

\begin{table}\label{tab:eBeam}
    \centering
    \caption{ESRF and EBS main electron beam parameters. Electron beam parameters for the high-$\beta$ (ESRF) and the EBS straight sections. Values taken at the symmetry point of the straight section, where the insertion devices are placed. Values fir ESRF are taken from \cite{ESRF2014}. For EBS, two (very close) design values are used, from Ref.~\cite{ESRF2014} labelled EBS-S28A, and those coming from further lattice refinement, labelled EBS-S28D. \todo{Remove ESRF-AT?? Check again} }
        \resizebox{\textwidth}{!}{%
        \begin{tabular}{ccccc}
         Lattice & $\sigma_e$ horizontal {[}$\mu$m{]} &  $\sigma_e'$ horizontal {[}$\mu$rad{]} & $\sigma_e$ vertical {[}$\mu$m{]} & $\sigma_e'$ vertical {[}$\mu$rad{]} \\
         \input{tableElectron.txt}

@@

\section{Simulations}

@@

The source size is usually approximated supposing a Gaussian distribution of the electron sizes (or divergence) (Table \ref{tab:eBeam}) and undulator natural source size or divergence. Although the former is true, the latter is not. Let $\sigma_{e(h,v)}$ and $\sigma_{e(h,v)}'$ be the electron beam size and divergence at a centre of the symmetry of the straight section. The subscripts $h$ and $v$ stand for horizontal and vertical, respectively. The undulator natural source size and divergence are given by \cite{elleaume}:
\begin{align}
    \label{eq:photon small sigmas}
    \sigma_u=\frac{2.74}{4\pi}\sqrt{\lambda L}\approx \sqrt{\frac{\lambda L}{2 \pi^2}}  && \sigma_u' = 0.69\sqrt{\frac{\lambda}{L}}\approx \sqrt{\frac{\lambda}{2 L}},
\end{align}
where $\lambda$ is the emitted wavelength and $L=N\lambda_u$ is the undulator length, with $\lambda_u$ as the magnetic period of the undulator and $N$, the number of magnetic periods. The electron beam usually presents different horizontal and vertical beam emittances, but the undulator natural source size or divergence presents radial symmetry at the central cone (excluding polarization). The photon source size and divergence are given by the convolution of the electron sizes and the undulator natural sizes. If we consider Gaussian distributions, the photon source is
\begin{align}
\label{eq:photon big sigmas}
\Sigma_{h,v}=\sqrt{\sigma_{e(h,v)}^2 + \sigma_u^2} && \Sigma_{h,v}'=\sqrt{\sigma_{e(h,v)}'^2 + \sigma_u'^2}.
\end{align}

% $\Sigma$ is given by the convolution of the electron beam size and the undulator natural source size or equivalently, single electron source size \cite{elleaume}: a storage ring with low emittance will produce small photon source size. It is also convenient to define two other relationships: the first one relates the magnetic field $B$ in an undulator to a dimensionless parameter $K$, the deflection parameter: $K = 93.4\lambda_u[\text{m}]\text{B}[\text{T}]$. The second relationship shows the wavelengths at which the emitted spectral intensity os peaked, or resonance. It is a function of $K$; the harmonic number n; the Lorentz factor $\gamma$ and the observation angle $\phi$:
% \begin{align}
%     \lambda=\frac{\lambda_u}{2\text{n}\gamma^2}\Big(1+\frac{K^2}{2}+\gamma^2\phi^2\Big).
% \end{align}
% The first term is the relativistic contracted wavelength, the second term is the magnetic tuning and the last term is the off-axis wavelength increase \cite{elleaume}.

The beam sizes at different positions at the beamline were estimated in a different way depending if the beam is focused or not. If the beam is focused, its dimensions in a first approximation are the photon source size multiplied by the corresponding magnification factor. For predicting the photon beam in a non-focal position the beam divergence is used. In an ideal {\'{e}}tendue preserving optical system, where the conservation of the Smith-Helmholz invariante applies, the resulting beam divergence transforms inversely as the beam size: it can be calculated as the product between the beam divergence at the source with the demagnification factor $M^{-1}$.

Using the values for U18.3, $n$=1, and  $E=17225$ eV (Table \ref{tab:Undulators}), one obtains a focal size 139$\times$5 nm$^2$ (H$\times$V) for ESRF and 10$\times$5 nm$^2$ for EBS (Table~\ref{tab:HandCalculations}). The focal spot for ESRF lattice is too large in horizontal for the beamline specification, therefore the VSS could be closed to 50$\mu$m thus reducing the beamsize at the VSS plane by a factor ca. 8 \inred{$\approx$ 403.8/50=8.076}. Therefore the horizontal size at the sample will be affected by the same factor obtaining roughly 17$\times$5 nm$^2$

In general, for grazing incident optics with large demagnifications ($M^{-1} \gg 100$) imaging extended sources, the effect of optical aberrations is very important even using a perfect optical surface that focus point-to-point. As a consequence, the results calculated using only the demagnification factor are usually too optimistic. Moreover imperfections in the optical surface (figure errors, slope errors and microroughness) are often the limiting factors of the real reflective optics. Their effect will studied in Sections~\ref{level1} and \ref{level2}. The smearing of the focal size because of diffractive effects by the mirror clipping will also be discussed.

@@

The next step would be to estimate the number of photons at the sample. For that, we need the spectrum at the source. It is not straightforward to get this information analytically but one can calculate the flux and power using different codes. Perhaps the most performant and well maintained codes spread in the synchrotron community are SPECTRA \cite{codeSPECTRA} and SRW \cite{codeSRW}. The XOPPY package in the OASYS suite implements three codes: US\cite{codeUS}, URGENT: \cite{codeURGENT} and SRW \cite{codeSRW}. Figure~\ref{fig:FluxU18} shows the spectra calculated over an aperture if 1 $\times$ 1 mm$^2$ placed at $d$=27.1 m. This aperture opening is set to fully accept the central cone at the resonance peak of the harmonic in use ($\max(\Sigma') \times d = 0.73$ mm), with $\Sigma'$ values from in Table~\ref{tab:HandCalculations}. The peak intensity is about  $2.9\cdot10^{14}$ photons/s/0.1{\%}bw for EBS and 1.9 10$^{14}$ for ESRF (see \ref{fig:FluxU18}). One can appreciate how the emittance plays an important role in both peak width and specially in peak amplitude. The number of photons per second after the multilayer monochromator is 2.9 10$^{14}$ for EBS and 1.9 10$^{14}$ for ESRF, considering that the energy bandwidth transmitted by a multilater monochromator is $\Delta E/E \approx 10^{-2}$ and the bandwidth in \ref{fig:FluxU18} is 10$^{-3}$, and supposing at this point ideal reflectivity (100\%) for the multilayers.

@@

Next we would like to estimate the geometrical transmitivity of the beamline, therefore we have to consider which optical element crop the beam and remove photons. This can be analysed using their angular acceptance or numerical aperture. The multilayer monochromator accepts the whole beam. We consider a VSS closed to $S_h=50~\mu$m in horizontal and fully opened in vertical. For ESRF its transmittivity is 1/8 (as discussed before) and 1 for EBS. In a similar way, the KB mirrors will crop the beam to a size that is the projection of their length to a direction perpendicular to the optical axis. Using mirror length of $L_v$=60 mm and $L_h$=26 mm for VFM and HFM, respectively, and incident angle of 15 mrad, the KB is behaving like an aperture of $900 \times 390 \mu m^2$. We get N.A. values of 6.2$\times$21.6 $mu$rad (aperture size divided by the distance from the mirror to the VSS in H and to the source in V). The transmittivity coefficient is the ratio of the accepted N.A. over the beam divergence, resulting 9.5\% (H) and 0.18\% (V)for ESRF and 16.3\% and 0.18\% for EBS. The overall transmission is the product of H and V ratios, therefore 0.2\% for ESRF and 2.95\% for EBS, therefore 3.8 $10^{11}$ photons/s for ESRF and 8.5 $10^{12}$ photons/s for EBS. 
\todo{check with references $10^9$ at 33.6 keV. I get: 1.6e+15 * 0.00295 * 0.7**3 = 1.62e+12. How closed are the slits? }

One can also discuss some simple consequences about the beam coherence. The coherence fraction (CF) \cite{arxivCF} of the undulator beam is the occupation of the lowest coherent mode, as it will be discussed in Section~\ref{level2}. An approximated estimation of it can be obtained using Eqs.~\ref{eq:photon small sigmas} and \ref{eq:photon big sigmas}. The CF in its the horizontal and vertical planes are
\begin{equation}\label{eq:coherent fraction}
 CF_{(h,v)} = \frac{\sigma_u \sigma'_u}{\Sigma_{(h,v)} \Sigma'_{(h,v)}},
\end{equation}
and the total coherent fraction is the product of both (CF=CF$_h \times $CF$_v$). An estimation of the CF calculated in such a mode give CF=0.12\% for the ESRF and 2.73\% for EBS at 17 keV photon energy. The CF roughly indicates the ratio of the ``coherent photons'' over the total number of photons. It is a good indicator of the quality of a coherent beam (a fully coherent beam has CF=1 and an incoherent beam CF=0). The beamline optics acts as a ``coherence filter'' by removing the ``bad'' or ``incoherent'' photons and in consequence increasing the CF. A complete cleaning is impossible because of the spatial overlapping of the different coherence modes. However, apertures centered at the optical axis would make the work. If we compare the values transmittivity of the beamline due to geometrical clipping  (0.2\% for ESRF and 2.95\% for EBS) with the the CF values, we remark that they are very close. This implies that the beam at the focal position would be highly coherent because the fraction of the photons removed from the beam matches the CF value that quantify the fraction of ``good'' or ``coherent'' photons.

To assess the focused beam size, one can idealize the beamline in two parts. First, the ML monochromator and VSS that will prepare a coherent plane wave as the source is very far from the sample position. Second, the KB has a double role, acting as a focusing element of focal distance $f$=5 cm, but also crops the beam as as it would be done by a square aperture of $D \times D=$ 400$\times$400 $\mu m^2$ (the length of the mirror projected in a plane perpendicular to the beam propagation) \inred{(we obtained 900 in H, that will be reduced by 0.5 as propagated from 10 cm to 5 cm upstrem from the sample, and 390 in V approximated by 400)}. This aperture has two effects: i) as discussed before, it removes a lot of off-axis photons increasing the CF to a value close than one, and ii) make a broadening of the focal spot because of coherent diffraction. We know from optics that the diffraction by a square aperture of a collapsing spherical wavefront has an intensity distribution proportional to $\sinc^2(k D x/2 f) \sinc^2(k D y / 2 f)$, where $k=2\pi/\lambda$, and $x$ ($y$) is the horizontal (vertical) coordinate . Considering that the FWHM of $\sinc(x)$ is approximately 2.78, one obtains a FWHM of 7.75$\times$7.75 nm$^2$ for the focal spot. This spot size is of the order or larger than the one obtained by geometrical considerations, therefore contributing significantly to the size of the focused beam. One can conclude that for the conditions in use this is a diffraction limited beamline.

In summary, we show in this section how simple analytical calculations that can be done in its major part ``by hand'' help to estimate the geometrical beam size at focal position (17$\times$5 nm$^2$ for ESRF, 10$\times$5 nm$^2$ for the EBS) the flux (3.8 $10^{11}$ photons/s for ESRF and 8.5 $10^{12}$ photons/s for EBS) and also anticipate that the beam is highly coherent producing a diffraction limited focal spot of  7.75$\times$7.75 nm$^2$. We ignored at this point the effect of real optics containing surface errors, the reflectivity of the elements, and other aspects that will be treated in depth in the next sections. This is, however, the first step to be done when analyzing a beamline. It also helps in having a critical eye or sanity check on the numerical calcuations that will provide more precised values, but must remain the this order of magnitude.

>>>>

\begin{abstract}
We consider different approaches to simulate a modern X-ray beamline. Several methodologies with increasing complexity are applied to discuss the relevant parameters that quantify the beamline performances. Parameters such as flux, dimensions and intensity distribution of the focused beam and coherence properties are obtained from simple analytical calculations to sophisticated computer simulations using ray tracing and waveoptics techniques. A last generation x-ray nanofocusing beamline for coherent  applications (ID16A at the ESRF) has been chosen  to study in detail the issues related to highly demagnifying sychrotron sources and exploiting the beam coherence. The performance of the beamline is studied for two storages rings: the old ESRF (emittance 4000 pm) and the new ESRF-EBS (emittance 150 pm). In addition to traditional results in term of flux and beam sizes, an innovative study on the partial coherence properties based on propagation of coherent modes is presented. The different algorithms and methodologies are implemented in the software suite OASYS \cite{codeOASYS}. They are discussed with emphasis in benefits and limitations of each one. 
\end{abstract}

@@

Many storage ring based x-ray synchrotron facilities are building or planning upgrades to increase brilliance and coherent flux by one to three orders of magnitude.  The first upgrade of a large facility will be the EBS (Extremely Brilliant Source) \cite{orangebook} at the European Synchrotron Radiation Facility (ESRF). The new storage ring of 150 pm emittance that will be commissioned in 2020 will significantly boost the x-ray coherence of the synchrotron beams. Applications exploiting beam coherence such as x-ray photon correlation spectroscopy, coherent diffraction imaging, propagation-based phase contrast imaging and ptychography will strongly benefit from this update.

Accurate calculation and quantitative evaluation of the parameters related to x-ray optics is of paramount importance for designing, building and exploiting the new beamlines. Every modern beamline follows a procedure of conceptual design plus technical design where detailed simulations are essential. These calculations permit to test the design parameters in a virtual computer environment where advantages and limitations can be studied quantitatively. The selection to the best performant optics configuration requires the detailed simulation of the optics imperfections, that in many cases are the limiting factor of the beamline optical configuration. Deformation in the optical elements due to heat load need to be controlled, therefore optics simulations have to include results from engineering modeling of the thermal deformation associated to mechanical issues usually dome using finite eelement analysis. Traditionally the optics calculations for synchrotron beamlines are done using ray tracing techniques. They model the x-ray beam as a collection of different rays that are incoherent one to another. This technique is still essential for obtaining information on aberrations, flux propagation, monochromator bandwidth, etc. However, the high coherence of the new sources make necessary to complement the ray tracing methods with models including effects of diffraction and scattering from a coherent beam. A fully coherent beam is still a dream for the new generation storage rings. In spite of a high increate of the coherent fraction of the new beams, the coherent fraction at tipically hard x-ray photon energies is still at few percent for the new sources. This makes still important a ray tracing analysis that does not become obsolete, but has to be complemented with an analysis based on wave optics. Because of the non perfect coherence the wave optics analysis becomes non trivial and can be done with different levels of approximation, as discussed in this work. In addition, wave optics simulations are usually expensive from the computer point of view.

For illustrating the applicability of a battery of methods and solutions for beamline optics simulations we have chosen the ESRF beamline ID16A. This beamline provides a high-brilliance beam focused to a few nanometer. It shows characteristics increasingly searched in modern beamlines: extremely high demagnification (sub-micrometer beam) and high coherence. It combines coherent imaging techniques and X-ray fluorescence microscopy to perform quantitative 3D characterization of the morphology and the elemental composition of specimens in their native state.

X-ray beamlines are very particular optical systems. Some implications of exploiting the low wavelengths such as the necessity of using grazing incidence reflectors and the use of perfect crystals as typical hard x-ray monochromators make useless, or at least incomplete, the application of common commercial software packages for the study of synchrotron beamlines. For this reason many optics tools have been developed by the synchrotron community. Some of them have a long history like SHADOW (more than 40 years old) or SRW (almost 30), for mentioning only codes used in this work. These historical tools together with other new applications have been agglutinated in a user friendly software OASYS. Many of the calculations presented here have been performed directly in OASYS or with script created with the help of OASYS.

The paper is structured as follows: we first introduce in Section~\ref{Beamline description} the main parameters of the beamline and make the selection of the particular choice of the parameters used in the calculations. Section~\ref{Simulations} is the central part of the paper, and describe the different methods used to analyse the beamline with increasing complexity, setting a hierarchy of methodologies that can be exported to the simulations of any beamline. In Section \ref{level0} we calculate with simple methods approximated values of focal size, flux and coherent fraction. Then accuracy is improved using geometrical methods in Section \ref{level1} that give reasonable results as far as some corrections are considered because of the  mentioned characteristics of high demagnification and high coherence of the studied beamline. The wave optics section~\ref{level2} also include different models whose complexity increases when the results expected are more and mode demanding. The simulations methods are presented in a hierarchical approach:  they provide more and better results but the price to pay is to deal with more complex and computationally expensive tools.

 %-------------------------------------------------------------------------
 %
 %-------------------------------------------------------------------------

@@

\resizebox{\textwidth}{!}{%
\begin{tabular}{lcccccc}
name  & $\lambda_u$ [mm] & $L$ [m] & N & $K$ (11200 eV) & $K$ (17225 eV) & $K$ (33600 eV) \\ 
\input{tableUndulator.txt}
\end{tabular}%
}
\end{table}

@@

\begin{table}\label{tab:Positions}
    \caption{Position of the main ID16A optical elements with respect to the source (undulator centre). It also shows the values for horizontal and vertical demagnification $M^{-1}=p/q$, where $p$ and $q$ are the the distances object-optics and optics-image, respectively.}
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{rcccccc}
    -                      & Source & Multilayer & Slit  & KB (v) & KB (h) & Focal plane \\ 
    Position {[}m{]}:      & 0.00   & 28.30      & 40.00 & 184.90 & 184.95 & 185.00      \\
    Horizontal $M^{-1}$:   & -      & 2.42       & -     & -      & 2899   & -           \\
    Vertical $M^{-1}$:     & -      & -          & -     & 1849   & -      & -           \\ 
    \end{tabular}%
    }%
\end{table}

Simulations were done using the two different storage ring lattices mentioned before: the (already dismantled) ESRF DBA where the undulators are in a high-$\beta$ straight section (larger horizontal emittance) and the new EBS (lower horizontal emittance) straight sections. The parameters for the current and the new lattice are shown in Tables \ref{tab:TableSources} and \ref{tab:eBeam}.

@@

\begin{table}\label{tab:eBeam}
    \centering
    \caption{ESRF and EBS main electron beam parameters. Electron beam parameters for the high-$\beta$ (ESRF) and the EBS straight sections. Values taken at the symmetry point of the straight section, where the insertion devices are placed. Values for ESRF are taken from \cite{ESRF2014}. For EBS, two (very close) design values are used, from Ref.~\cite{ESRF2014} labelled EBS-S28A, and those coming from further lattice refinement, labelled EBS-S28D. \todo{Remove ESRF-AT and EBS-S28A} }
        \resizebox{\textwidth}{!}{%
        \begin{tabular}{ccccc}
         Lattice & $\sigma_e$ horizontal {[}$\mu$m{]} &  $\sigma_e'$ horizontal {[}$\mu$rad{]} & $\sigma_e$ vertical {[}$\mu$m{]} & $\sigma_e'$ vertical {[}$\mu$rad{]} \\
         \input{tableElectron.txt}

@@

\section{Simulations}
\label{Simulations}

@@

The source size is usually approximated supposing a Gaussian distribution of the electron sizes (or divergence) (Table \ref{tab:eBeam}) and undulator natural source size or divergence. Fo some cases, as the undulator emission, the Gaussian approximation may not be accurate and have some implications that will be discussed later. Let $\sigma_{e(h,v)}$ and $\sigma_{e(h,v)}'$ be the electron beam size and divergence standard deviation of the electron distributions at a centre of the symmetry of the straight section in a plane perpendicular to the main propagation direction. The subscripts $h$ and $v$ stand for horizontal and vertical, respectively.

The undulator natural source size and divergence are given by \cite{elleaume}:
\begin{align}
    \label{eq:photon small sigmas}
    \sigma_u=\frac{2.74}{4\pi}\sqrt{\lambda L}\approx \sqrt{\frac{\lambda L}{2 \pi^2}}  && \sigma_u' = 0.69\sqrt{\frac{\lambda}{L}}\approx \sqrt{\frac{\lambda}{2 L}},
\end{align}
where $\lambda$ is the emitted wavelength and $L=N\Lambda_u$ is the undulator length, with $\Lambda_u$ as the magnetic period of the undulator and $N$, the number of magnetic periods. The electron beam usually presents different horizontal and vertical beam emittances, but the undulator natural source size or divergence presents radial symmetry at the central cone (excluding polarization). The photon source size and divergence are approximately given by the convolution of the electron sizes and the undulator natural sizes. If we consider Gaussian distributions, the photon source is
\begin{align}
\label{eq:photon big sigmas}
\Sigma_{h,v}=\sqrt{\sigma_{e(h,v)}^2 + \sigma_u^2} && \Sigma_{h,v}'=\sqrt{\sigma_{e(h,v)}'^2 + \sigma_u'^2}.
\end{align}

The beam sizes at different positions at the beamline were estimated in a different way depending if the beam is focused or not. If the beam is focused, its dimensions in a first approximation are the photon source size multiplied by the corresponding magnification factor. For predicting the photon beam sizes in a position out of the focus the beam divergence is used. In an ideal {\'{e}}tendue preserving optical system, where the conservation of the Smith-Helmholz invariante applies, the resulting beam divergence transforms inversely as the beam size: it can be calculated as the product between the beam divergence at the source with the demagnification factor $M^{-1}$.

Using the values for U18.3, $n$=1, and  $E=17225$ eV (Table \ref{tab:Undulators}), one obtains a focal size 139$\times$5 nm$^2$ (H$\times$V) for ESRF and 10$\times$5 nm$^2$ for EBS (Table~\ref{tab:HandCalculations}). The focal spot for ESRF lattice is too large in horizontal for the beamline specification, therefore the VSS could be closed to 50$\mu$m thus reducing the beamsize at the VSS plane by a factor ca. 8 \inred{$\approx$ 403.8/50=8.076}. Therefore, for ESRF, the horizontal size at the sample will be affected by the same factor obtaining roughly 17$\times$5 nm$^2$

In general, for grazing incident optics imaging extended sources with large demagnifications ($M^{-1} \gg 100$), the effect of optical aberrations is very important even using a perfect optical surface that focus point-to-point. As a consequence, the results calculated using only the demagnification factor are usually too optimistic. Moreover imperfections in the optical surface (figure errors, slope errors and microroughness) are often the limiting factors of the real reflective optics. Their effect will studied in Sections~\ref{level1} and \ref{level2}.
%The smearing of the focal size because of diffractive effects by the mirror clipping will also be discussed.

@@

Next, we want to estimate the number of photons at the sample. For that, we need the spectrum at the source. It is not straightforward to get this information analytically but one can calculate the flux and power using different codes. Perhaps the most performant and well maintained codes spread in the synchrotron community are SPECTRA \cite{codeSPECTRA} and SRW \cite{codeSRW}. The XOPPY package in the OASYS suite implements three codes: US \cite{codeUS}, URGENT \cite{codeURGENT} and SRW. Figure~\ref{fig:FluxU18} shows the spectra calculated over an aperture if 1 $\times$ 1 mm$^2$ placed at $d$=27.1 m. This aperture opening is set to fully accept the central cone at the resonance peak of the harmonic in use ($\max(\Sigma') \times d = 0.73$ mm), with $\Sigma'$ values from in Table~\ref{tab:HandCalculations}. The peak intensity is about  $2.9\cdot10^{14}$ photons/s/0.1{\%}bw for EBS and 1.9 10$^{14}$ for ESRF (see \ref{fig:FluxU18}). One can appreciate how the emittance plays an important role in both peak width and specially in peak value. The number of photons per second after the multilayer monochromator is 2.9 10$^{15}$ for EBS and 1.9 10$^{15}$ for ESRF, considering that the energy bandwidth transmitted by a multilater monochromator is $\Delta E/E \approx 10^{-2}$ and the bandwidth in Fig.~\ref{fig:FluxU18} is 10$^{-3}$, and supposing at this point ideal reflectivity (100\%) for the multilayers.

@@

To estimate the geometrical transmitivity of the beamline we have to consider which optical elements crop the beam and remove photons. This can be analysed using their angular acceptance or numerical aperture. The multilayer monochromator accepts the whole beam. We consider a VSS closed to $S_h=50~\mu$m in horizontal and fully opened in vertical. For ESRF its transmittivity is 1/8 (as discussed before) and 1 for EBS. In a similar way, the KB mirrors will crop the beam to a size that is the projection of their length to a direction perpendicular to the optical axis. Using mirror length of $L_v$=60 mm and $L_h$=26 mm for VFM and HFM, respectively, and incident angle of 15 mrad, the KB is behaving like an aperture of $900 \times 390 \mu m^2$. We get N.A. values of 6.2$\times$21.6 $mu$rad (aperture size divided by the distance from the mirror to the VSS in H and to the source in V). The transmittivity coefficient is the ratio of the accepted N.A. over the beam divergence, resulting 9.5\% (H) and 0.18\% (V)for ESRF and 16.3\% and 0.18\% for EBS. The overall transmission is the product of H and V ratios, therefore 0.2\% for ESRF and 2.95\% for EBS, therefore 3.8 $10^{12}$ photons/s for ESRF and 8.5 $10^{13}$ photons/s for EBS. 
\todo{check with references $10^9$ at 33.6 keV. I get: 1.6e+15 * 0.00295 * 0.7**3 = 1.62e+12. How closed are the slits? We used the undulator peak value over the whole ML bandwidth, that overestimates the integrated flux.}

One can also discuss some simple consequences about the beam coherence. The coherence fraction (CF) \cite{arxivCF} of the undulator beam is the occupation of the lowest coherent mode, as it will be discussed in Section~\ref{level2}. An approximated estimation of it can be obtained using Eqs.~\ref{eq:photon small sigmas} and \ref{eq:photon big sigmas}. The coherent fraction is the ratio of the phase space for the coherent beam (zero emittance) over the actual beam. The CF in its the horizontal and vertical planes are
\begin{equation}\label{eq:coherent fraction}
 CF_{(h,v)} = \frac{\sigma_u \sigma'_u}{\Sigma_{(h,v)} \Sigma'_{(h,v)}},
\end{equation}
and the total coherent fraction is the product of both (CF=CF$_h \times $CF$_v$). An estimation of the CF calculated in such a mode give CF=0.12\% for the ESRF and 2.73\% for EBS at 17 keV photon energy. The CF roughly indicates the ratio of the ``coherent photons'' over the total number of photons. It is a good indicator of the quality of a coherent beam (a fully coherent beam has CF=1 and an incoherent beam CF=0). The beamline optics acts as a ``coherence filter'' by removing the ``bad'' or ``incoherent'' photons and in consequence increasing the CF. A complete cleaning is impossible because of the spatial overlapping of the different coherence modes. However, apertures centered at the optical axis would clean the beam from these undesired photons. If we compare the values transmittivity of the beamline due to geometrical clipping  (0.2\% for ESRF and 2.95\% for EBS) with the the CF values, we remark that they are very close. This suggests that the beam at the focal position will be highly coherent because the fraction of the photons removed from the beam matches the CF value that quantify the fraction of ``good'' or ``coherent'' photons.

To assess the focused beam size given by a coherent beam, one can idealize the beamline in two parts. First, the ML monochromator and VSS that will prepare a coherent plane wave as the source is very far from the sample position. Second, the KB has a double role, acting as a focusing element of focal distance $f$=5 cm, but also crops the beam as as it would be done by a square aperture of $D \times D=$ 400$\times$400 $\mu m^2$ (the length of the mirror projected in a plane perpendicular to the beam propagation) 
\todo{REMOVE: we obtained 900 in H, that will be reduced by 0.5 as propagated from 10 cm to 5 cm upstrem from the sample, and 390 in V approximated by 400}. This aperture has two effects: i) as discussed before, it removes a lot of off-axis photons increasing the CF to a value close than one, and ii) make a broadening of the focal spot because of coherent diffraction. We know from optics that the diffraction by a square aperture of a collapsing spherical wavefront has an intensity distribution proportional to $\sinc^2(k D x/2 f) \sinc^2(k D y / 2 f)$, where $k=2\pi/\lambda$, and $x$ ($y$) is the horizontal (vertical) coordinate . Considering that the FWHM of $\sinc(x)$ is approximately 2.78, one obtains a FWHM of 7.75$\times$7.75 nm$^2$ for the focal spot. This spot size is of the order or larger than the one obtained by geometrical considerations, therefore contributing significantly to the size of the focused beam. One can conclude that for the conditions in use this is a diffraction limited beamline.
